Probability

### 2016-04-11 prob.tex
prob.tex [Quantum probability measure] updated....

\begin{example}{[}Quantum probability measure{]} Like in the classical
case, there are $2^{8}$ events. But as Mermin explains in a simple
example~\cite{MerminPRL1990}, here is a possible probability measure:
\[
\begin{array}{rcl}
\mathbb{P}_{xxx}(\{LLL\}) & = & \frac{1}{4}\\
\mathbb{P}_{xxx}(\{LLR\}) & = & 0\\
\mathbb{P}_{xxx}(\{LRL\}) & = & 0\\
\mathbb{P}_{xxx}(\{LRR\}) & = & \frac{1}{4}\\
\mathbb{P}_{xxx}(\{RLL\}) & = & 0\\
\mathbb{P}_{xxx}(\{RLR\}) & = & \frac{1}{4}\\
\mathbb{P}_{xxx}(\{RRL\}) & = & \frac{1}{4}\\
\mathbb{P}_{xxx}(\{RRR\}) & = & 0\\
\mathbb{P}_{xxx}(E) & = & \sum_{\omega\in E}\mathbb{P}_{xxx}(\left\{ \omega\right\} )
\end{array}
\]
In contrast with the classical example previously, the probability
measure of the space cannot be expressed as the product of the probability
measure of each electron, i.e., 
\[
\mathbb{P}_{xxx}(\{abc\})\ne\mathbb{P}(\{a\})\mathbb{P}(\{b\})\mathbb{P}(\{c\})\textrm{ .}
\]
Correlated variables are also common in the classical world. For example,
if the first coin is head or tail decided by whether the temperature
is higher than a particular degree and the second coin is decided
by whether Coca Cola is sold more than a particular amount, then we
know these two coins are correlated. Einstein, Podolsky, and Rosen
(EPR)~\cite{EPR1935} suggested any correlated quantum probability
results may be interpreted classically. For example, the nature might
roll an tetrahedron die with $\{LLL,LRR,RLR,RRL\}$ in its four faces.
If $LRR$ were face-down, the measurement result for three electrons
would be $L$, $R$, and $R$, respectively. Furthermore, this tetrahedron
die might not be accessed by human for some reasons so that we think
we are handling a system constituting three electrons.

In order to verify EPR's idea, we need to analysis the idea of probability
more carefully. Imagining somebody tosses a coin behind a veil, we
heard the coin is tossed, but cannot see the result. Although the
probability for us is still $\mathbb{P}(\{H\})=\mathbb{P}(\{T\})=\frac{1}{2}$,
we know the coin behind the veil must be either head or tail face-up.

In classical probability, there is no difference between the coins
have not been tossed or the coins have been tossed but we do not know.
However, there is a difference when we consider the quantum probability.
Because the three electrons can be specially separated, and each electron
can be measured along the $x$ axis separately, if the nature actually
rolled the tetrahedron die, this die should be rolled before the electrons
are separated and measured. In another word, there should be a period
that the human observer would not know the measurement results for
each electron along the $x$ axis, but the nature should know. The
result for $j$-th electron is denoted by $w\left(\sigma_{x}^{j}\right)$
so that $w\left(\sigma_{x}^{1}\right)w\left(\sigma_{x}^{2}\right)w\left(\sigma_{x}^{3}\right)\in\{LLL,LRR,RLR,RRL\}$,
i.e., the number of $L$ in $w\left(\sigma_{x}^{1}\right)$, $w\left(\sigma_{x}^{2}\right)$,
and $w\left(\sigma_{x}^{3}\right)$ should be odd.

The interesting part is that three electrons cannot be measured the
spin only along the $x$ axis, but also along the $y$ axis with the
result down ($D$) or up ($U$). We only consider to measure even
number of electrons along the $y$ axis, and the measure of non-singleton
set should be computed by $\mathbb{P}_{ijk}(E)=\sum_{\omega\in E}\mathbb{P}_{ijk}(\left\{ \omega\right\} )$.
\begin{eqnarray*}
\begin{array}{rcl}
\mathbb{P}_{xyy}(\{LDD\}) & = & 0\\
\mathbb{P}_{xyy}(\{LDU\}) & = & \frac{1}{4}\\
\mathbb{P}_{xyy}(\{LUD\}) & = & \frac{1}{4}\\
\mathbb{P}_{xyy}(\{LUU\}) & = & 0\\
\mathbb{P}_{xyy}(\{RDD\}) & = & \frac{1}{4}\\
\mathbb{P}_{xyy}(\{RDU\}) & = & 0\\
\mathbb{P}_{xyy}(\{RUD\}) & = & 0\\
\mathbb{P}_{xyy}(\{RUU\}) & = & \frac{1}{4}
\end{array} & \begin{array}{rcl}
\mathbb{P}_{yxy}(\{DLD\}) & = & 0\\
\mathbb{P}_{yxy}(\{DLU\}) & = & \frac{1}{4}\\
\mathbb{P}_{yxy}(\{DRD\}) & = & \frac{1}{4}\\
\mathbb{P}_{yxy}(\{DRU\}) & = & 0\\
\mathbb{P}_{yxy}(\{ULD\}) & = & \frac{1}{4}\\
\mathbb{P}_{yxy}(\{ULU\}) & = & 0\\
\mathbb{P}_{yxy}(\{URD\}) & = & 0\\
\mathbb{P}_{yxy}(\{URU\}) & = & \frac{1}{4}
\end{array} & \begin{array}{rcl}
\mathbb{P}_{yyx}(\{DDL\}) & = & 0\\
\mathbb{P}_{yyx}(\{DDR\}) & = & \frac{1}{4}\\
\mathbb{P}_{yyx}(\{DUL\}) & = & \frac{1}{4}\\
\mathbb{P}_{yyx}(\{DUR\}) & = & 0\\
\mathbb{P}_{yyx}(\{UDL\}) & = & \frac{1}{4}\\
\mathbb{P}_{yyx}(\{UDR\}) & = & 0\\
\mathbb{P}_{yyx}(\{UUL\}) & = & 0\\
\mathbb{P}_{yyx}(\{UUR\}) & = & \frac{1}{4}
\end{array}
\end{eqnarray*}
Similarly, there would be predetermined $w\left(\sigma_{x}^{j}\right)$
and $w\left(\sigma_{y}^{j}\right)$ for these probabilities. The number
of $L$ or $D$ in $\left\{ w\left(\sigma_{x}^{1}\right),w\left(\sigma_{y}^{2}\right),w\left(\sigma_{y}^{3}\right)\right\} $,
$\left\{ w\left(\sigma_{y}^{1}\right),w\left(\sigma_{x}^{2}\right),w\left(\sigma_{y}^{3}\right)\right\} $,
and $\left\{ w\left(\sigma_{y}^{1}\right),w\left(\sigma_{y}^{2}\right),w\left(\sigma_{x}^{3}\right)\right\} $
should be even. If we look these 9 letters carefully, we can find
every $w\left(\sigma_{x}^{j}\right)$ appear once and every $w\left(\sigma_{y}^{j}\right)$
appear twice. Hence, the number of $L$ in $w\left(\sigma_{x}^{1}\right)$,
$w\left(\sigma_{x}^{2}\right)$, and $w\left(\sigma_{x}^{3}\right)$
should be even. This contradict to the conclusion in our last paragraph.
Therefore, EPR's assumption is wrong and the correlation of quantum
probability cannot be interpreted by a classical probability. \end{example}

>>>>

\begin{example}{[}Quantum probability measure{]} Like in the classical
case, there are $2^{8}$ events. But as Mermin explains in a simple
example~\cite{MerminPRL1990}, here is a possible probability measure:
\[
\begin{array}{rcl}
\mathbb{P}_{xxx}(\{LLL\}) & = & \frac{1}{4}\\
\mathbb{P}_{xxx}(\{LLR\}) & = & 0\\
\mathbb{P}_{xxx}(\{LRL\}) & = & 0\\
\mathbb{P}_{xxx}(\{LRR\}) & = & \frac{1}{4}\\
\mathbb{P}_{xxx}(\{RLL\}) & = & 0\\
\mathbb{P}_{xxx}(\{RLR\}) & = & \frac{1}{4}\\
\mathbb{P}_{xxx}(\{RRL\}) & = & \frac{1}{4}\\
\mathbb{P}_{xxx}(\{RRR\}) & = & 0\\
\mathbb{P}_{xxx}(E) & = & \sum_{\omega\in E}\mathbb{P}_{xxx}(\left\{ \omega\right\} )
\end{array}
\]
In contrast with the classical example previously, the event of different
electrons are not independent or correlated. More precisely, consider
the event for each electron separately:
\begin{eqnarray*}
F_{1,L} & = & \left\{ LLL,LLR,LRL,LRR\right\} \\
F_{2,L} & = & \left\{ LLL,LLR,RLL,RLR\right\} \\
F_{3,L} & = & \left\{ LLL,LRL,RLL,RRL\right\} 
\end{eqnarray*}
They are correlated means 
\[
\mathbb{P}_{xxx}(F_{1,L}\cap F_{2,L}\cap F_{3,L})=\mathbb{P}_{xxx}(\{LLL\})=\frac{1}{4}\ne\frac{1}{8}=\mathbb{P}_{xxx}(F_{1,L})\mathbb{P}_{xxx}(F_{2,L})\mathbb{P}_{xxx}(F_{3,L})\textrm{ .}
\]

Classical events may also be correlated even they come from different
sources. A typical example is events defined by the temperature is
highly likely correlated to the event related to how Coca Cola is
sold. Another example are the same events we discussed in coin-tossing.
However, this time the coins are tossed behind a veil, and somebody
tosses the three coins for you. Because she tosses the coins behind
a veil, she might actually roll an tetrahedron die with $\{HHH,HTT,THT,TTH\}$
in its four faces. If $HTT$ were face-down, she places three coins
with $H$, $T$, and $T$ face-up, respectively. Then, she uncovers
the veil, and claims she has tossed the coins. Then, the result of
coin-tossing is correlated, and we will never see $TTT$ if the coins
are tossed in this way.

Because we do not know how nature decides the spin of an electron,
Einstein, Podolsky, and Rosen (EPR)~\cite{EPR1935} suggested the
nature might give us the probability measure~$\mathbb{P}_{xxx}$
because she rolled an tetrahedron die or performed other classical
and deterministic process behind the veil. This claim is convincing
if $\mathbb{P}_{xxx}$ is the only probability measure we have, but
will lead to a contradiction if we consider other probability measures
all together. Notice that after the coins are placed before uncovering
the veil, which face of a coin is up is decided although we do not
know. This would be also true for the quantum probability measure.
Because the three electrons can be specially separated, and each electron
can be measured along the $x$ axis separately, if the nature rolled
the tetrahedron die, this die should be rolled before the electrons
are separated and measured, and she should know the result of measurement.
Let the result of the $j$-th electron measured along the $x$ axis
be $w\left(\sigma_{x}^{j}\right)$ so that $w\left(\sigma_{x}^{1}\right)w\left(\sigma_{x}^{2}\right)w\left(\sigma_{x}^{3}\right)\in\{LLL,LRR,RLR,RRL\}$,
i.e., the number of $L$ in $w\left(\sigma_{x}^{1}\right)$, $w\left(\sigma_{x}^{2}\right)$,
and $w\left(\sigma_{x}^{3}\right)$ should be odd.

The interesting part is that three electrons cannot be measured the
spin only along the $x$ axis, but also along the $y$ axis with the
result down ($D$) or up ($U$). We only consider to measure even
number of electrons along the $y$ axis, and the probability measure
of non-singleton set should be computed by $\mathbb{P}_{ijk}(E)=\sum_{\omega\in E}\mathbb{P}_{ijk}(\left\{ \omega\right\} )$.
\begin{eqnarray*}
\begin{array}{rcl}
\mathbb{P}_{xyy}(\{LDD\}) & = & 0\\
\mathbb{P}_{xyy}(\{LDU\}) & = & \frac{1}{4}\\
\mathbb{P}_{xyy}(\{LUD\}) & = & \frac{1}{4}\\
\mathbb{P}_{xyy}(\{LUU\}) & = & 0\\
\mathbb{P}_{xyy}(\{RDD\}) & = & \frac{1}{4}\\
\mathbb{P}_{xyy}(\{RDU\}) & = & 0\\
\mathbb{P}_{xyy}(\{RUD\}) & = & 0\\
\mathbb{P}_{xyy}(\{RUU\}) & = & \frac{1}{4}
\end{array} & \begin{array}{rcl}
\mathbb{P}_{yxy}(\{DLD\}) & = & 0\\
\mathbb{P}_{yxy}(\{DLU\}) & = & \frac{1}{4}\\
\mathbb{P}_{yxy}(\{DRD\}) & = & \frac{1}{4}\\
\mathbb{P}_{yxy}(\{DRU\}) & = & 0\\
\mathbb{P}_{yxy}(\{ULD\}) & = & \frac{1}{4}\\
\mathbb{P}_{yxy}(\{ULU\}) & = & 0\\
\mathbb{P}_{yxy}(\{URD\}) & = & 0\\
\mathbb{P}_{yxy}(\{URU\}) & = & \frac{1}{4}
\end{array} & \begin{array}{rcl}
\mathbb{P}_{yyx}(\{DDL\}) & = & 0\\
\mathbb{P}_{yyx}(\{DDR\}) & = & \frac{1}{4}\\
\mathbb{P}_{yyx}(\{DUL\}) & = & \frac{1}{4}\\
\mathbb{P}_{yyx}(\{DUR\}) & = & 0\\
\mathbb{P}_{yyx}(\{UDL\}) & = & \frac{1}{4}\\
\mathbb{P}_{yyx}(\{UDR\}) & = & 0\\
\mathbb{P}_{yyx}(\{UUL\}) & = & 0\\
\mathbb{P}_{yyx}(\{UUR\}) & = & \frac{1}{4}
\end{array}
\end{eqnarray*}
Similarly, the nature should predetermine $w\left(\sigma_{x}^{j}\right)$
and $w\left(\sigma_{y}^{j}\right)$ for these probabilities. Furthermore,
because she do not know along which axis we are going to measure,
she should predetermine the same $w\left(\sigma_{x}^{j}\right)$ and
$w\left(\sigma_{y}^{j}\right)$ for all different probabilities. Notice
that the number of $L$ or $D$ in $\left\{ w\left(\sigma_{x}^{1}\right),w\left(\sigma_{y}^{2}\right),w\left(\sigma_{y}^{3}\right)\right\} $,
$\left\{ w\left(\sigma_{y}^{1}\right),w\left(\sigma_{x}^{2}\right),w\left(\sigma_{y}^{3}\right)\right\} $,
and $\left\{ w\left(\sigma_{y}^{1}\right),w\left(\sigma_{y}^{2}\right),w\left(\sigma_{x}^{3}\right)\right\} $
should be even. If we look these 9 letters carefully, we can find
every $w\left(\sigma_{x}^{j}\right)$ appear once and every $w\left(\sigma_{y}^{j}\right)$
appear twice. Hence, the number of $L$ in $w\left(\sigma_{x}^{1}\right)$,
$w\left(\sigma_{x}^{2}\right)$, and $w\left(\sigma_{x}^{3}\right)$
should be even. This contradict to the conclusion in our last paragraph.
Therefore, EPR's assumption is wrong, and it is not always true that
the nature can predetermine the measurement result before we perform
the measurement. \end{example}
